---
title: "Unit2 Assignments"
author: "Sri Seshadri"
date: "10/29/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy = T, out.width = 60)
```

#Chapter 3 Question 1

```{r}

Male_Yes <- 10
Male_No <- 30
Female_Yes <- 6
Female_No <- 34

# Probabilities of use among male and females
p.male.use <- Male_Yes/(Male_Yes + Male_No)
p.female.use <- Female_Yes/(Female_Yes + Female_No)
# Odds among male and females
odds_Male_use <- Male_Yes/Male_No
odds_Female_use <- Female_Yes/Female_No

# odds ratio Male vs Female

OR.Male.to.Female <- odds_Male_use/odds_Female_use

Q1df <- data.frame(Prob.Male.use = p.male.use, Prob.Female.use = p.female.use, Odd.Male.use = odds_Male_use, Odds.Female.use = odds_Female_use, `OddsRatio m/f` = OR.Male.to.Female)

knitr::kable(round(Q1df,2))
```


# Chapter 3 Question 2

```{r}
religion <- read.csv(file = 'religion.csv')
#sanitycheck <- do.call(rbind,dfapply(religion,favstats, select = is.numeric))
#knitr::kable(round(sanitycheck,2), caption = "Summary statistics")

# note there is missing data
attending.respondents <- sum(religion$RELSCHOL)
Total.respondents <- nrow(religion)

odds_relschol <- attending.respondents/(Total.respondents - attending.respondents)
prob_relschol <- attending.respondents/(Total.respondents)

ct <- descr::CrossTable(religion$RELSCHOL,religion$RACE,prop.r = F,prop.c = F,prop.chisq = F,prop.t = F)
ct
crosstable <- table(religion$RELSCHOL,religion$RACE)
Prob.nonwhite.relschool.attend <- crosstable[2,1]/sum(crosstable[,1])
Prob.white.relschool.attend <- crosstable[2,2] / sum(crosstable[,2])

odds.nonwhite.relschool.attend <- crosstable[2,1]/crosstable[1,1]
odds.white.relschool.attend <- crosstable[2,2]/crosstable[1,2]
OR <- odds.white.relschool.attend/odds.nonwhite.relschool.attend

```



a. The odds of attending religious school is 0.1465201
    and the probability is 0.1277955
b. 
    Probability of non-white attending religious school = `r Prob.nonwhite.relschool.attend`
    Probability of white attending religious school = `r Prob.white.relschool.attend`
    
    odds of non-white attending religious school = `r odds.nonwhite.relschool.attend`
    odds of white attending religious school = `r odds.white.relschool.attend`
    Odds ratio of WHITE over NON-WHITE = `r OR`
    

# Chapter 3 Question 3

```{r}
model1 <- glm(RELSCHOL ~ RACE, family = binomial,data = religion)
model2 <- glm(RELSCHOL ~ RACE + ATTEND + INCOME, family = binomial,data = religion)

#summary(model1)
coef1 <- coefficients(model1)[2]
odds.ratio <- exp(coef1)

AIC_model1 <- AIC(model1)
BIC_model1 <- BIC(model1)

AIC_model2 <- AIC(model2)
BIC_model2 <- BIC(model2)

q3bdf <- data.frame(Model = c(1,2),AIC = c(AIC_model1,AIC_model2), BIC = c(BIC_model1,BIC_model2))
```

a. The odds ratio of white and non-white attendance based on logistic regression model is `r as.character(odds.ratio)`. It agrees with the question 2b above.

b. 
```{r}
knitr::kable(q3bdf, caption = "AIC & BIC stats")
```

    based on the numbers above model 2 would be chosen. The lower the A/BICs the better the model.
    
c. 
```{r}
source("getequation.R")
eqn <- getequation("RELSCHOL",model2)
RACE <- c(1,0)
ATTEND <- c(5,5)
INCOME <- c(4,4)

logoddsRELSHOL <- -3.58 - 1.29 * RACE  + 0.33 *  ATTEND  + 0.2 *  INCOME
odds_RELSCHOL <- round(exp(logoddsRELSHOL),2)

paste("The odds of attendance for white and non white are ", round(odds_RELSCHOL[1],2), "and",round(odds_RELSCHOL[2],2), "respectively")
```

d. 
```{r}
# summary(model1)
# summary(model2)

adj.odds.Ratio <- exp(coefficients(model2)[2])

#paste("The adjusted odds ratio for race is ", adj.odds.Ratio)
paste(" When effects of income and attend are controlled, change in race has an multiplicative effect of ", adj.odds.Ratio)
```
The adjusted odds ratio for race is `r adj.odds.Ratio`
When effects of income and attend are controlled, change in race has an multiplicative effect of ` r adj.odds.Ratio`


# Chapter 3 Question 4

```{r}
model1.probit <- glm(RELSCHOL ~ RACE, family = binomial(link = 'probit'),data = religion)
model2.probit <- glm(RELSCHOL ~ RACE + ATTEND + INCOME, family = binomial(link = 'probit'),data = religion)

prob.White.attendance <- pnorm(sum(coefficients(model1.probit)))
prob.NonWhite.attendance <- pnorm(coefficients(model1.probit)[1])
```

a. The prob of white and non white attendances based on probit are `r prob.White.attendance` and `r prob.NonWhite.attendance` respectively. They compare well to 2b.

b. 
```{r}
AIC_model1.probit <- AIC(model1.probit)
BIC_model1.probit <- BIC(model1.probit)

AIC_model2.probit <- AIC(model2.probit)
BIC_model2.probit <- BIC(model2.probit)

q4bdf <- data.frame(Model = c(1,2),AIC = c(AIC_model1.probit,AIC_model2.probit), BIC = c(BIC_model1.probit,BIC_model2.probit))

knitr::kable(q4bdf, caption = "AIC and BICs for probit")
```

Model 2 is still the winner, lower the A/BIcs the better the model.

c. 
```{r}
eqn2 <- getequation("ATTEND",model2.probit)
RACE <- c(1,0)
ATTEND <- c(5,5)
INCOME <- c(4,4)
predicted.probs <- pnorm(-2.07 - 0.73 * RACE  + 0.19 *  ATTEND  + 0.12 *  INCOME)
```

The predicted probability for white is `r round(predicted.probs[1],2)` and for non white `r round(predicted.probs[2],2)`

```{r}
RACE <- c(0,0)
ATTEND <- c(4,4)
INCOME <- c(4,10)
predicted.probs <- pnorm(-2.07 - 0.73 * RACE  + 0.19 *  ATTEND  + 0.12 *  INCOME)
change.prob <- predicted.probs[2] - predicted.probs[1]

```

d. The discrete change in predicted probability is `r change.prob`

# Chapter 3 Question 5

Non-whites have lower probability of attending religious school, while income and service attendance increases the probability of religious attendance.
