---
title: "Wine sales prediction"
author: "Sri Seshadri"
date: "11/11/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

# 1. Introduction

A large wine manufacturer is studying data collected on 12,000 commercially available wines, with a goal to predict the number of cases ordered based upon the characteristics. The manufacturer intends to adjust the winde offerings based on the findings. The data collected is related to chemical properties of wine with response variable being the number of sample cases sold to distribution companies. the data also includes features like review stars provided by the tasters and label appeal. 

# 2. Exploratory Data analysis

The below table shows the summary statistics of the data. It is seen that there are missing data. "Stars" has the most missing values. The missing values mostly correspond to NO (0) cases sold. Which is most likely due to lack of opportunity to sample because of none sold.

```{r}
winedata <- read.csv(file = 'wine.csv')
library(mosaic)
# Compute overall statistic and show missing values
sanitycheck <- do.call(rbind,dfapply(winedata,favstats, select = is.numeric))
rowname <- rownames(sanitycheck)

rownames(sanitycheck) <- rowname
knitr::kable(round(sanitycheck,2), caption = "Summary Stats and missing values")
Missing.Stars <- as.factor(ifelse(is.na(winedata$STARS),1,0))
Missing.Alch <- as.factor(ifelse(is.na(winedata$Alcohol),1,0))
Missing.Sulph <- as.factor(ifelse(is.na(winedata$Sulphates),1,0))
Missing.pH <- as.factor(ifelse(is.na(winedata$Sulphates),1,0))
Missing.TSD <- as.factor(ifelse(is.na(winedata$TotalSulfurDioxide),1,0))
Missing.FSD <- as.factor(ifelse(is.na(winedata$FreeSulfurDioxide),1,0))
Missing.Chl <- as.factor(ifelse(is.na(winedata$Chlorides),1,0))
Missing.ResSug <- as.factor(ifelse(is.na(winedata$ResidualSugar),1,0))

winedataeda <- cbind(winedata, Missing.ResSug,Missing.Chl,Missing.FSD,Missing.TSD,Missing.pH,Missing.Sulph,Missing.Alch,Missing.Stars)

# Write to file for examination

# write.csv(file = 'winedatamod.csv', winedata)

```


```{r, fig.cap="Missing STARS values' association with Number of cases sold", fig.height=4}
library(ggplot2)
p <- ggplot(data = winedataeda) 
p + geom_bar(mapping = aes(x = TARGET, col = Missing.Stars, fill = Missing.Stars)) + theme_classic()

#winedata %>% group_by(TARGET) %>% summarise_at("STARS", function(x) sum(is.na(x))/length(x)*100)
```

Figure 2 shows the historgram of features in the data. The chemical properties of wines appear to share an identical distribution with peaks closer to zero. This mau be likely to some standardization done to the day. 
The variable "TARGET" looks to be poisson or negative biniomially distributed with inflation at 0. 


```{r,fig.cap="Histograms of features"}
windedataMelt <- reshape2::melt(winedata[,-1])
ggplot(data = windedataMelt,mapping = aes(x = value)) + geom_histogram() + facet_wrap(~variable,scales = "free_x") + theme_bw()
```

\pagebreak

# 3. Feature selection

In this section we'll attempt to select important features that explain the target variable. We'll explore the correlations that might exist in the data. There is positive correlations between TARGET and STARS and LabelAppeal. 

```{r, fig.cap="Correlation plot"}
wine.complete <- winedata[complete.cases(winedata),]
cordata <- cor(wine.complete)
corrplot::corrplot(cordata, tl.cex = 0.7)

winedata <- winedata %>% mutate(TARGET0 = as.factor(ifelse(TARGET > 0 ,0, 1)))
```

It'll be useful to identify what contributes to the zero inflation in the TARGET variable. For which let's create an indicator variable "TARGET0" to equal 1 when TARGET is 0 and 0 otherwise. 

## 3.1 Decision Trees

### 3.1.1 Predictors for zero cases sold (No sale)

Figure 4 shows the decision tree for TARGET0. Where 1 is no sale (cases sold = 0) and 0 is sale (cases sold > 0 ). Figure 5 shows the variable importance plot after a random forest bootstrap. While LabelAppeal did not contribute to node purity (Decrease in Gini index), it did affect accuracy on out of bag (OOB) samples. 

```{r, fig.cap= "Decision tree - Sale (0) or no Sale (1)"}
predictors <- colnames(winedata)[c(-1,-2,-17)]
response <- c("TARGET","TARGET0")
tr <- rpart::rpart(formula = TARGET0 ~ ., data = winedata[,c(predictors, response[2])], method = "class")
library(partykit)
plot(partykit::as.party(tr), gp = gpar(fontsize = 8 ))
```

```{r, fig.cap= "Variable Importance - Randomforest"}
randtr <- randomForest::randomForest(formula = TARGET0 ~ ., data = winedata[complete.cases(winedata),c(predictors, response[2])], importance = T)
randomForest::varImpPlot(randtr,main = 'Variable Importance plot - Random forest')
```


### 3.1.2 Predictors for cases sold; when successfuly sold (cases > 0)

Figure 6 shows the decision tree of cases sold when they are greater than 0. It can be seen that the LabelAppeal and STARS are the top hitters. Figure 7 shows the variable importance plot when a random forest method is employed, where a random set of predictors are chosen at each iteration to fit a decision tree. Variables whose exclusion contributes to higher Mean Squared Error (MSE) is deemed important. LabelApeal, STARS and Alcohol are top 3 variables that are important.

```{r, fig.cap="Simple tree for TARGET"}
windedata.filtered <- winedata %>% filter(TARGET0 == 0)
tr2 <- rpart::rpart(formula = TARGET ~ ., data = windedata.filtered[,c(predictors, response[1])])
plot(as.party(tr2),gp = gpar(fontsize = 8 ))
```


```{r, fig.cap="Variable importance plot - TARGET" }
randtr2 <- randomForest::randomForest(formula = TARGET ~ ., data = windedata.filtered[complete.cases(windedata.filtered),c(predictors, response[1])], importance = T)
randomForest::varImpPlot(randtr2, main = "Variable importance plot from random forest - TARGET")
```

\pagebreak

# 4. Training and validation samples

The data is split into training and validation sample for evaluating models for final selection. The 75% of the data is sampled as training sample and the rest is set aside as validation set. Figure 8 shows that the TARGET variable is identically distributed for both training and validation samples.

```{r, fig.cap="TARGET distribution for training and validation samples"}
set.seed(10)
train <- winedata %>%  dplyr::sample_frac(0.75)
test <- winedata[!winedata$INDEX %in% train$INDEX,]
par(mfrow = c(1,2))
hist(train$TARGET, main = "TARGET from training sample", xlab = "TARGET", col = "red")
hist(test$TARGET, main = "TARGET from test sample", xlab = "TARGET", col = "blue")


```


# 5. Modeling

The following modeling approaches will be tried for predicting the number of cases sold.

* Poisson regression
* Negative binomial regression
* Zero infalted Poisson regression
* Zero inflated Negative binomial regression
* OLS regression

## 5.1 Poisson regression.

As seen in table 1, the mean and standard deviation of TARGET are not too far apart. Poisson model is a good candidate for regression modeling.  

### 5.1.1 Simple model with Label appeal as the predictor.

While we know that the TARGET variable is zero inflated and there are atleast more than one important predictor from the section above, we'll attempt to build the model ground up with LabelAppeal as the single predictor. 

```{r}
poisson.mdl1 <- glm(TARGET ~ LabelAppeal,family = poisson(link = 'log'), data = train)
summary(poisson.mdl1)
lmtest::lrtest(poisson.mdl1)
glance.poisson.mdl1 <- broom::glance(poisson.mdl1)
glance.poisson.mdl1$pseudoR.Sq <- (poisson.mdl1$null.deviance - poisson.mdl1$deviance)/poisson.mdl1$null.deviance
poisson.mdl1.predict <- predict(poisson.mdl1,newdata = subset(test,!is.na(test$LabelAppeal)), type = 'response')
poisson.mdl1.coefs <- broom::tidy(poisson.mdl1)
glance.poisson.mdl1$MAE <- mean(abs(test$TARGET[!is.na(test$TARGET)] - round(poisson.mdl1.predict,0)))
knitr::kable(glance.poisson.mdl1,digits = 2,caption = "Simple Poisson model statistics")


# poisson.mdl2 <- glm(TARGET ~ LabelAppeal + STARS + Alcohol,family = poisson(link = 'log'), data = train)
# summary(poisson.mdl2)
# 
# poisson.full <- glm(TARGET ~ .,family = poisson(link = 'log'), data = train[,c(-1,-17)] )
# summary(poisson.full)
# 1 -pchisq(deviance(poisson.mdl2)-deviance(poisson.full),11)

```

#### 5.1.1.1 Model interpretation

It can be seen from the Likelihood ratio test ab0ve, that the slope of LabelAppeal is NOT zero and an unit increase in label appeal increases by `r round(exp(poisson.mdl1.coefs[2,2]),0)`. However the model is not an adequate fit with residual deviance over degrees of freedom being much higher than 1 (in comparison to a saturated model). This is also reflected in the Pseudo R Squared value. 

The model does not predict 0s. As can be see from figure 9. 

```{r, fig.cap= "Simple poisson regression diagnostics",fig.height=6}
layout <- matrix(c(1,2,3,4,5,5), ncol = 2, byrow = T)
layout(layout)
plot(poisson.mdl1)

plot(poisson.mdl1$data$TARGET , round(poisson.mdl1$fitted.values,0), xlab = "TARGET", ylab = "Rounded fitted")
```

### 5.1.2 Poisson model with Label appeal & Stars as predictors

Now, we'll use Label appeal and stars as predictors and compare the model with the simple model in the above section.

```{r}

poisson.mdl2 <- glm(TARGET ~ LabelAppeal + STARS, family = poisson(link = "log"), data = train[!is.na(train$LabelAppeal),])
summary(poisson.mdl2)
lmtest::lrtest(poisson.mdl2)
plot(train$TARGET[!is.na(train$STARS)], round(poisson.mdl2$fitted.values,0))
```


```{r, eval=F}
predictors <- colnames(winedata)[unlist(mosaic::dfapply(winedata, is.numeric))][-1]
poisson.mdl <- glm(TARGET ~ STARS + LabelAppeal + Alcohol, family = poisson(link = 'log'),data = winedata[complete.cases(winedata),])
summary(poisson.mdl)

poisson.mld.full <- glm(TARGET ~ ., family = poisson(link = 'log'), data = winedata[complete.cases(winedata),predictors])
summary(poisson.mld.full)

poisson.mdl2 <- glm(TARGET ~ STARS + LabelAppeal, family = poisson(link = 'log'),data = winedata[complete.cases(winedata),])

lmtest::lrtest(poisson.mdl, poisson.mdl2)

plot(poisson.mld.full, col = winedata[complete.cases(winedata),'TARGET'])
hist(poisson.mdl$fitted.values)
hist(poisson.mdl2$fitted.values)

nb.mdl <- MASS::glm.nb(TARGET ~ STARS + LabelAppeal + Alcohol,data = winedata[complete.cases(winedata),])
summary(nb.mdl)

models <- list(M1 = poisson.mdl, M2 = poisson.mdl2, M3 = poisson.mld.full, M4 = nb.mdl)
sapply(models,glance)

lmtest::lrtest(poisson.mld.full,nb.mdl)
plot(winedata[complete.cases(winedata),"TARGET"],round(fitted(poisson.mdl),0))


```

